---
title: Spatial Regression - Mississippi <br><small>Advanced Data Analytics</small></br>
author: Nate Selleck
output:
  html_document: 
    df_print: paged
    rows.print: 10
    theme: cosmo
    highlight: breezedark
    number_sections: yes
    toc: no
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document: default
  html_notebook:
    df_print: paged
    rows.print: 10
    theme: cosmo
    highlight: breezedark
    number_sections: yes
    toc:  no
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options:
  chunk_output_type: inline
  mode: gfm
---
<body style = "background-color:#b1d3aa">

```{r Packages, echo=TRUE, message=FALSE, warning=FALSE}

pacman::p_load("biscale", "cleangeo", "cowplot", "dplyr", "geosphere", "ggplot2", "maps", "maptools", "rgdal", "rgeos", "sf", "sp", "spatialreg", "spdep", "tidyr", "mapproj")

```

# Read in the Data

```{r Data Chunk}

data <- read.csv('./Data/childpov18_southfull.csv', 
                 colClasses = c("character", "character", "character", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric"))

names(data)[names(data)=="X2016.child.poverty"] <- "child.pov.2016"

ms_pov <- data %>% subset(State == "MS")

summary(ms_pov)

```

# Run the OLS

```{r Little Bit of OLS}

equation <- child.pov.2016 ~ rural + urban + lnmanufacturing + lnag + 
  lnretail + lnhealthss + lnconstruction + lnlesshs + 
  lnunemployment + lnsinglemom + lnblack + lnhispanic + 
  lnuninsured + lnincome_ratio + lnteenbirth + lnunmarried

options(scipen = 5)

ols <- lm(equation, data=ms_pov)
summary(ols)

```

# Contiguity Neighbors

```{r Howdy Contiguity Neighbors}

fips <- county.fips
fips.codes <- separate(data = fips, col = polyname, into = c("state", "county"), sep = ",")
ms_fips <- subset(fips.codes, state=="mississippi", select=fips)

mississippi <- map(database = "county", regions = "mississippi", fill=T, plot=F)
ms_sp = map2SpatialPolygons(mississippi,ms_fips$fips,CRS("+proj=longlat"))

```

# Moran and LaGrange

```{r Who you callin a Moran}

cleaned <- clgeo_Clean(ms_sp)
neighb.data <- poly2nb(cleaned, queen=T)
cont.neighb <- nb2listw(neighb.data,style="W", zero.policy = TRUE)

lm.morantest(ols, cont.neighb)

```

$~$

Based on the p-value from the Moran Test, I can reject the null hypothesis that there is no spatial correlation in the residuals. Because there is spatial dependency in child poverty rates in Mississippi, I ran a LaGrange to find the best model

$~$

```{r LaGrange}

lm.LMtests(ols, cont.neighb, test="all")

```

$~$

The robust LM test for error dependence had the lowest p-value, so I created a Spatial Error model for the data.

$~$

# Spatial Error Model
```{r Spatial Error}

sp.err.model <- spatialreg::errorsarlm(equation, data=ms_pov, cont.neighb)
summary(sp.err.model, Nagelkerke = TRUE)

```

# K-nearest Neighbors

```{r K Neighbors}

all.xy <-centroid(ms_sp)
colnames(all.xy) <- c("x","y")

all.dist.k1 <- knn2nb(knearneigh(all.xy, k=1, longlat = TRUE))
all.dist.k3 <- knn2nb(knearneigh(all.xy, k=3, longlat = TRUE))
all.dist.k5 <- knn2nb(knearneigh(all.xy, k=5, longlat = TRUE))

#Determine max k distance value to neighbor
all.max.k1 <- max(unlist(nbdists(all.dist.k1, all.xy, longlat=TRUE)))
all.max.k3 <- max(unlist(nbdists(all.dist.k3, all.xy, longlat=TRUE)))
all.max.k5 <- max(unlist(nbdists(all.dist.k5, all.xy, longlat=TRUE)))

#Calculate neighbors based on distance
all.sp.dist.k1 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k1, longlat = TRUE)
all.sp.dist.k3 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k3, longlat = TRUE)
all.sp.dist.k5 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k5, longlat = TRUE)

#Create neighbor list
all.dist.neighb.k1 <- nb2listw(all.sp.dist.k1,style="W", zero.policy = TRUE)
all.dist.neighb.k3 <- nb2listw(all.sp.dist.k3,style="W", zero.policy = TRUE)
all.dist.neighb.k5 <- nb2listw(all.sp.dist.k5,style="W", zero.policy = TRUE)

```

```{r Distance Lag}

all.dist.lag.k1 <- spatialreg::lagsarlm(equation, data = ms_pov, listw = all.dist.neighb.k1)
all.dist.lag.k3 <- spatialreg::lagsarlm(equation, data = ms_pov, listw = all.dist.neighb.k3)
all.dist.lag.k5 <- spatialreg::lagsarlm(equation, data = ms_pov, listw = all.dist.neighb.k5)

summary(all.dist.lag.k1, NagelKerke = TRUE)

```

```{r Distance Error}

all.dist.err.k1 <- spatialreg::errorsarlm(equation, data = ms_pov, listw = all.dist.neighb.k1)
all.dist.err.k3 <- spatialreg::errorsarlm(equation, data = ms_pov, listw = all.dist.neighb.k3)
all.dist.err.k5 <- spatialreg::errorsarlm(equation, data = ms_pov, listw = all.dist.neighb.k5)

summary(all.dist.err.k1, Nagelkerke = TRUE)

```
$~$

Based on the p-values for the Distance Lag and Distance Error models, the distance to the nearest neighbor does not seem to affect child poverty rates within a county.

$~$

# Time to Make Maps

```{r MAP}

dist.err.data <- summary(sp.err.model, correlation=TRUE, Nagelkerke = TRUE)

dist.err.output <- cbind.data.frame(ms_pov$FIPS,
                                    dist.err.data$fitted.values, 
                                    dist.err.data$residual, 
                                    ms_pov$child.pov.2016, 
                                    ms_pov$urban, 
                                    ms_pov$lnunemployment, 
                                    ms_pov$lnlesshs, 
                                    ms_pov$lnsinglemom,
                                    stringsAsFactors = FALSE)

#Renaming columns
colnames(dist.err.output) <- c("fips","fitted","resid","childpov",
                        "urban","unemployment","less_hs","single_mom")

```

```{r Urban Map, fig.height=6, fig.width=8}

ms_fortify <- fortify(ms_sp)

ms_poly <- merge(x = ms_fortify, y = dist.err.output, 
                 by.x = "id", by.y = "fips", all = TRUE)

urban_data <- bi_class(ms_poly, x = childpov, y = urban, 
                           dim = 3, style = "quantile")

legend_urban <- bi_legend(pal = "DkViolet",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Urbanization",
                    size = 6)

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))

urban_pov_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = urban_data, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "DkViolet", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-94,-85), ylim=c(29,36)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Urbanization") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))

final_urban_map <- ggdraw() +
  draw_plot(urban_pov_map, 0, 0, 1, 1) +
  draw_plot(legend_urban, 0.60, 0.035, 0.3, 0.3)
final_urban_map

```

```{r Unemployment, fig.height=6, fig.width=8}


unemployment_data <- bi_class(ms_poly, x = childpov, y = unemployment, 
                           dim = 3, style = "quantile")

legend_unemp <- bi_legend(pal = "DkViolet",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Unemployment",
                    size = 6)

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))

unemp_pov_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = unemployment_data, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "DkViolet", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-94,-85), ylim=c(29,36)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Unemployment Rate") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))


final_unemp_map <- ggdraw() +
  draw_plot(unemp_pov_map, 0, 0, 1, 1) +
  draw_plot(legend_unemp, 0.60, 0.035, 0.3, 0.3)
final_unemp_map

```

```{r Less HS, fig.height=6, fig.width=8}


lesshs_data <- bi_class(ms_poly, x = childpov, y = less_hs, 
                           dim = 3, style = "quantile")

legend_hs <- bi_legend(pal = "DkViolet",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Less HS",
                    size = 6)

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))

hs_pov_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = lesshs_data, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "DkViolet", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-94,-85), ylim=c(29,36)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and High School Graduates") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))


final_hs_map <- ggdraw() +
  draw_plot(hs_pov_map, 0, 0, 1, 1) +
  draw_plot(legend_hs, 0.60, 0.035, 0.3, 0.3)
final_hs_map

```

```{r Single Mom, fig.height=6, fig.width=8}


singlemom_data <- bi_class(ms_poly, x = childpov, y = single_mom, 
                           dim = 3, style = "quantile")

legend_mom <- bi_legend(pal = "DkViolet",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Single Mother\n Households",
                    size = 6)

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))

mom_pov_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = singlemom_data, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "DkViolet", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-94,-85), ylim=c(29,36)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Single Mother Households") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))


final_mom_map <- ggdraw() +
  draw_plot(mom_pov_map, 0, 0, 1, 1) +
  draw_plot(legend_mom, 0.60, 0.035, 0.3, 0.3)
final_mom_map


```

